---
title: "Travelling_sales_person"
author: "Pankaj Shah"
date: "6/15/2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Library 
```{r message = FALSE}
library(tidyverse)  # dplyr, ggplot
library(ggmap)      # for geom_leg() call
library(RCurl)      # for reading csv from github
library(dbscan)     # clustering
library(geosphere)  # calculates distances
library(TSP)        # solves travelling salesperson problem
library(gridExtra)  # combines ggplot charts
```

# Read Medical School Data:

```{r echo = T}
Locations <- 
    getURL("https://raw.githubusercontent.com/shahnp/data/master/Medschools.txt") %>%
    read.csv(text=., stringsAsFactors = F) %>%  # the dot means "this data" in the pipe
    as.data.frame()     

#identify the following variables
    Locations$ID    <- Locations$UnitID # the unique identifier for each site
    Locations$lon   <- Locations$lon        # Y coord
    Locations$lat   <- Locations$lat        # X coord
    Locations$value <- Locations$Enrollment  # size of points, weight for clustering
   
#What does this data describe
data_title <- "Medical Schools by Enrollment (2015)"
    
#Want to use groups? 
    #If yes, write "TRUE" and assign a group. 
    #If no, write "FALSE" and then select clustering method

  use_group <- TRUE
  Locations$group <- Locations$Region #column to group by geographically 

#clsutering method    
  clustMethod <- c("ward.D", "ward.D2", "single", "complete",
                   "average", "mcquitty", "median", "centroid")
  which_clustMethod <- 1 #defaults to 1: hclust(method = "ward.D")

```


```{r}
head(Locations)
```

# What if user_group == F

```{r}
if(use_group == F){
    
    clusters <- hclust(dist(select(Locations, lat, lon)), 
                       method = clustMethod[which_clustMethod])

    #summary(clusters$height)
    Locations$group <- paste0("cluster ",
                              cutree(clusters, h = mean(clusters$height)))
    
    cluster_bar <- 
        Locations %>% 
        group_by(group) %>% 
        summarise(nPoints = n()) %>% 
        arrange(-nPoints) %>% 
        ungroup() %>%
        mutate(Ord = row_number())
    
    ggplot(cluster_bar) +
      geom_boxplot(aes(y=nPoints, x = 1)) +
      coord_flip() +
      ggtitle(paste0("Distribution of points among the ",
                     nrow(cluster_bar), " clusters")) +
      theme(axis.text.y = element_blank(),
            axis.title.y = element_blank(),
            aspect.ratio = .25)
    

} else {
    Locations$group <- Locations$group
}
```


# Airport data
```{r}
Airports <- getURL("https://raw.githubusercontent.com/shahnp/data/master/airports.txt") %>%
                  read.csv(text=., stringsAsFactors = F) %>%  
                    as.data.frame()        
```


```{r}
head(Airports)
```

Create centers by finding the centroids of each group, then find the closest site to the centroid, then calculate distances to that central location

```{r}
Regions <- 
    Locations %>%
    select(group) %>%
    distinct() %>% 
    arrange(group)

Centroid <- 
    Locations %>%
    select(ID, group, lon, lat) %>% 
    group_by(group) %>% 
        mutate(lon.C = median(lon),
               lat.C = median(lat)) %>% 
        ungroup() %>% 
  
    #get distances to centroid of region
    rowwise() %>% 
        mutate(dist.mi = distGeo(p1 = c(lon, lat), 
                                 p2 = c(lon.C, lat.C))/1609.34) %>%
        ungroup() %>% 
    #identify central facility
    group_by(group) %>% 
        arrange(dist.mi) %>% 
        mutate(Ord = row_number()) %>% 
        ungroup() %>%
    mutate(Central = ifelse(Ord == 1, "Central", "Other")) %>% 
    #assign central facility as navigation point (lon.C, lat.C)
    group_by(group) %>% 
        arrange(Ord) %>% 
        mutate(lon.C = ifelse(Ord == 1, lon, 0),
               lat.C = ifelse(Ord == 1, lat, 0)) %>%
        mutate(lon.C = min(lon.C),
               lat.C = max(lat.C)) %>% 
        ungroup() %>% 
    #get distances for each site to central facility
    rowwise() %>% 
        mutate(dist.mi = distGeo(p1 = c(lon, lat), 
                                 p2 = c(lon.C, lat.C))/1609.34) %>% 
                                 #convert from meters to miles
        ungroup()
```


Generate map elements

```{r}
#map elements that repeat across maps    

map_airports <- geom_point(data = Airports, aes(x = lon, y = lat), 
                             size = 3, alpha = .5, color = "black", 
                             fill = "yellow", shape = 24)
  
map_legend <- theme(legend.position = "bottom",
                      axis.text = element_blank(),
                      axis.ticks = element_blank(),
                      axis.title = element_blank(),
                      panel.background = element_rect(fill = "white"))
  
all_states <- map_data("state")
  
all_counties <- map_data("county")

basemap <- 
      ggplot() +
      geom_polygon( data=all_counties, aes(x=long, y=lat, group = group),
                    colour="grey85", fill="GREY95") +
      geom_polygon( data=all_states, aes(x=long, y=lat, group = group),
                    colour="grey60", fill = NA, size = .5)
basemap
```

Generate polygon outlines

```{r}
group_outline <- data.frame(lon = vector("numeric", 0),
                            lat = vector("numeric", 0),
                            group = vector(class(Regions$group), length = 0))

for(i in 1:nrow(Regions)){
  group_points <-  Locations %>% 
                      filter(group == Regions$group[i]) %>% 
                      select(lon, lat)
  select_hull <- chull(group_points)
  select_hull <- c(select_hull, select_hull[1])
  
  group_outline <- rbind(group_outline,
                         group_points[select_hull, ] %>%
                          mutate(group = Regions$group[i]))

}

```


Show basemap

```{r}
# Make map

basemap +
geom_point(data = Locations,
            aes(lon, lat, color = factor(group), size = value), alpha = .8)+
geom_polygon(data = group_outline, 
              aes(lon, lat, group = group, 
                  fill = factor(group)), color = "grey50", alpha = .1) +
coord_map() +
theme(legend.position = "none", 
      axis.text = element_blank(),
      axis.ticks = element_blank(),
      axis.title = element_blank(),
      panel.background = element_rect(fill = "white")) +
ggtitle(paste0(data_title, ": ",
                nrow(Regions), " groups, n=", sum(Locations$value), " (uses ",
                ifelse(use_group == T, "original groups", "clustering algorithm"), ")"))


```

###Create maps
Create three maps within each group: 

* map showing region
* straight lines to central location
* shortest path through all points

Points are clustered within each region as a guide for the decision making process.

```{r eval = T, warning= FALSE}
#for loop to generate maps
for(i in 1:nrow(Regions)){

    RegionLoop <- 
        filter(Centroid, group == Regions$group[i]) %>%
        left_join(select(Locations, ID, value), by = "ID")
    
    #clusters
      clustersRL <- hclust(dist(select(RegionLoop, lat, lon)), 
                           method = clustMethod[which_clustMethod])

      RegionLoop$cluster <- cutree(clustersRL, h = mean(clustersRL$height))
      
      #table(RegionLoop$cluster)

    #creates shortest routes
      df <- select(RegionLoop, lon,lat)
      tsp <- TSP(dist(df))
      tour <- solve_TSP(tsp, method="nearest_insertion", control=list(rep=10))
      path.tsp <- unname(tour)
      RegionLoop$Path <- as.integer(tour)
  
    #plot(lat~lon, df[path.tsp,], type="b",asp=1)

    #map elements that repeat across maps    
      map_locations <- geom_point(data = RegionLoop, 
                                  aes(x = lon, y= lat, size = value, 
                                      fill = as.factor(cluster)), 
                                  alpha = .8, color = "black", shape = 21)
      
      map_location_central <- 
        geom_point(data = filter(RegionLoop, Central == "Central"),
                   aes(lon, lat),
                   color = "red")
      
      map_projection <- 
        coord_map(projection="mercator", 
                  xlim=c(min(RegionLoop$lon)-1, max(RegionLoop$lon)+1),
                  ylim=c(min(RegionLoop$lat)-1, max(RegionLoop$lat)+1))

    #show location of group    
    baseplot <- 
      ggplot() + coord_map() +
      geom_polygon(data=all_states, aes(x=long, y=lat, group = group),
                   color = NA, fill = "grey80", size = .5) +
       geom_point(data = RegionLoop, aes(x = lon, y= lat)) +
      #map_location_central +
      theme_nothing()

    #plots 1 & 2
    plot1 <- #map of distance to most central location
       basemap + map_projection + map_airports + #ggplot() +
       geom_leg(data = RegionLoop,
               aes(x = lon, xend = lon.C,
                   y = lat, yend = lat.C),
                color = "#E58700", size = 1, alpha = .5) +
        geom_leg(data = RegionLoop,
               aes(x = lon, xend = lon.C,
                   y = lat, yend = lat.C),
                color = "black", size = .5, alpha = .5) + 
        map_locations + map_location_central + map_legend +
        ggtitle(paste0(RegionLoop$group[1], 
                       ": ", nrow(RegionLoop), " sites (avg dist ",
                       round(median(RegionLoop$dist.mi), 0), " mi.)")) 
        
    plot2 <- #map of shortest routes
        basemap + map_projection + map_airports +  
        geom_path(data = RegionLoop[path.tsp,], 
                   aes(x=lon, y=lat), color = "#E58700", size = 1, alpha = .5) +
        geom_path(data = RegionLoop[path.tsp,], 
                   aes(x=lon, y=lat), color = "black", size = .5, alpha = .5) +  
        map_locations + map_location_central + map_legend +
        ggtitle(paste0("(n = ", sum(RegionLoop$value), ")"))

  #put all plots together
    grid.arrange(baseplot, plot1, plot2,ncol=3)

} #end of for loop

```


```{r}

```

```{r}

```

