---
title: "facebook_tweets"
author: "Pankaj Shah"
date: "6/20/2019"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Take tweet from twitter using rtweet package

- “Topic” : topic you will find at twitter 
-  n : how many tweet that you want to take 
-  include_rts : logical. If FALSE tweet taken didn’t contain retweet 
- lang = “en”  [English]

# Load the Library
```{r}
pacman::p_load("tm", "SnowballC", "wordcloud2", "RColorBrewer", "RCurl", "XML", "dplyr", "rtweet", "textclean")
```

# Authentication
```{r include=FALSE}
create_token(app = "xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx",
             consumer_key = "xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx",
             consumer_secret = "xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx",
             access_token = "xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx",
             access_secret = "xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx")

tweets <- search_tweets("facebook", n=500, include_rts = FALSE, lang = "en")
```

# Save into CSV
```{r}
# tweets_text <- tweets[,"text"] 
#   write.csv(tweets_text, "tweets.csv")
```

# Reading the save file
```{r}
tweets <- read.csv("https://raw.githubusercontent.com/shahnp/data/master/tweets.csv")
```


```{r}

# If we didn't save as csv then we can unnest tokens 

#Unnest the words - code via Tidy Text
# tweets_text  <- tweets_text  %>%
#                    tidytext::unnest_tokens(word, text)
```


# Duplicates
Lets Remove duplicate tweet using dplyr package
```{r}
tweets_text <- tweets %>% distinct(text)
```


# Check content text

```{r}
check_text(tweets_text$text)
```

# Stop_words

```{r, message=FALSE, error=FALSE}

#remove stop words - aka typically very common words such as "the", "of" etc
data(stop_words, package = "tidytext")
tweets_text  <- tweets_text  %>% filter(!text %in% stop_words$word)
```


# Empty Rows

Drop empty rows and NA text rows, then change class to character
```{r}
tweets_text <- tweets_text %>% 
                  drop_empty_row() %>% 
                    drop_NA()
tweets_text <- as.character(tweets$text)
```


# Replace contraction


```{r}
tweets_text <- replace_contraction(tweets_text) #wasn’t : was not, I’ll: I will, isn’t:is not , i’d : i would 

```

# Remove date

```{r}
tweets_text <- replace_date(tweets_text, replacement = "")
```

# Remove email


```{r}
tweets_text <- replace_email(tweets_text)
```

# Replace emoji

```{r}
tweets_text <- replace_emoji(tweets_text)
```


# Replace emoticon 
:) : smiley 
 :’( : crying 
  :* : kiss 
  
```{r}
tweets_text <- replace_emoticon(tweets_text)
```

# Replace grade

Replaces grades with word equivalents. Example: 
- A : very excellent 
- B+ : almost excellent 
- D : bad 
- etc

```{r}
tweets_text <- replace_grade(tweets_text)
```

# Remove hashtags

Replaces Twitter hash tags with blank character. Example: 
- #Rforbigdata : “” 
- #happynewyear2019 : “” 
- etc

```{r}
tweets_text <- replace_hash(tweets_text)
```

# Remove HTML

Remove angle braces and replaces HTML symbol markup with equivalent symbols. Example: 
- : “” 
- < &gt : < > 
- etc 
But, more recommended to add parameters in this function like symbol if you want to replace or remove HTML symbol markup. Example for argument symbol: 
symbol = TRUE/FALSE : logical. If FALSE HTML symbol markup will be removed

```{r}
tweets_text <- replace_html(tweets_text, symbol = FALSE)
```

# Remove incomplete sentence

Replaces incomplete sentence end marks with “|”. Example: 
- … : | 
- .? : | 
- etc 
But, more recommended to add parameters in this function like replacement if you don’t want to add | as replacement of incomplete sentence. Example for argument replacement: 
replacement = “” : replace incomplete sentence with blank character or don’t add |

```{r}
tweets_text <- replace_incomplete(tweets_text, replacement = "")
```

# Replace internet slang

Replaces internet slang with longer word equivalents that are more easily analyzed. Example: 
- 2nte : tonight 
- ASAP : as soon as possible 
- TGIF : thank god, it’s friday 
- etc

```{r}
tweets_text <- replace_internet_slang(tweets_text)
```

# Replace number

Numeric to be replaced with words. Example: 
- 1.997 : November sixteenth, one thousand nine hundred eighty 
- 28 : January thirty first, two thousand nineteen 
- 1,888 : - etc 
But, more recommended to add parameters in this function like replacement if you want to replace or remove date content. Example for argument replacement: 
replacement = “” : replace date content with blank character or remove date content

```{r}
tweets_text <- replace_number(tweets_text, remove = TRUE)
```


#Remove tag

Replaces Twitter style handle tags with blank character. Example: 
- @ramnath_vaidya : “” 
- @wulan123 : “” 
- @hadley : “” 
- etc

```{r}
tweets_text <- replace_tag(tweets_text)
```

#Remove url

Replace URLs with blank character. Example: 
- http://renkun.me/r/2014/07/26/difference-between-magrittr-and-pipeR.html : “” 
- ftp://cran.r-project.org/incoming/ : “” 
- etc

```{r}
tweets_text <- replace_url(tweets_text, replacement = "")
```

#Repalace word elongation

Replaces un-normalized forms with the most likely normalized form. Example: 
- noooooo! : no! - it’s sooo goooood : it’s so good - Wwwhhatttt! : what! - etc

```{r}
tweets_text <- replace_word_elongation(tweets_text)
```

Lets look at the head 
```{r}
tweets_text %>% head(5)
```


```{r}
# do a word count
# tweets_text %>%
#   count(word, sort = TRUE) %>% head(20) 
```


```{r}
# Remove other nonsense words
# tweets_text <-tweets_text %>%
#   filter(!word %in% c('t.co', 'https', 'handmaidstale', "handmaid's", 'season', 'episode', 'de', 'handmaidsonhulu',  'tvtime','watched', 'watching', 'watch', 'la', "it's", 'el', 'en', 'tv','je', 'ep', 'week', 'amp'))
```

#ˆWordCloud
WordCloud is used to display text data visually. By using WordCloud, the frequency of words can be displayed in an interesting form but still informative. More often a word is used, bigger the size of the word is displayed in WordCloud.

Function used for making WordCloud is wordcloud(). There are several parameter usually used in that function: 
- data : the text contain words that will be made wordCloud 
- min.freq : minimum frequency of words to be displayed 
- max.words : maximum words that will be displayed 
- random.order : words that have large frequencies will be plotted near the center 
- colors : coloring WordCloud

```{r}
library(tokenizers)
tweets_text <- tweets_text %>% 
                    tokenize_words() %>% 
                         as.character()

library(RColorBrewer)
library(wordcloud)
wordcloud(tweets_text,
         min.freq=1, 
         max.words = 80, 
         random.order = FALSE, 
         colors=brewer.pal(7, "Set2")) 
```
tdm : term-document matrix
freqTable : frequency table




